Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-01-02 Ben_Manuscripts/structure_paper/Draft.tex
Incorporated Gin comments

  Nanostructured membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control membrane architecture at the atomic scale
  permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

  Current commercial RO and NF membranes suffer limitations inherent to their
  fabrication. Although scalable, each process has a degree of stochasticity
  which makes overcoming the permeability-selectivity tradeoff a challenge.
  Namely, it is difficult to increase the permeability of a desired species,
  while maintaining the same retention of an undesired species
  \cite{werber_materials_2016}.

@@

  Commercial RO membranes are typically thin film composite membranes with a
  porous support layer and an active layer made of a polymer matrix formed
  through interfacial polymerization \cite{jeong_interfacial_2007}. During
  interfacial polymerization, a step-growth polymerization reaction occurs at the
  interface of an aqueous monomer solution, which has been soaked into the
  support layer, and an organic solution of a second monomer. The polymer matrix
  is dense and highly crosslinked which gives high rejections but low
  permeability.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically made by a phase inversion process  The most widely
  used phase inversion process is immersion precipitation, during which a polymer
  dissolved in a solvent is submerged in a non-solvent. A solid, porous membrane
  is all that remains once all solvent has been removed by non-solvent exchange
  \cite{smolders_microstructures_1992}.  Resultant pores are polydisperse in size
  which hurts membrane selectivity. A second technique used to create NF
  membranes is call track etching in which a polymer film is bombarded with
  charged particles, then chemically etched to create pores
  \cite{apel_track_2001}. The pores are uniform which benefits selectivity,
  however the membranes have a low porosity and subsequently low permeability.

@@

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-generation nanostructured
  membranes with high selectivity permitted by a precisely controlled pore size
  and high permeability allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Development of nanostructured materials has been limited by the ability to
  synthesize and scale various fundamentally sound technologies. Graphene sheets
  are atomically thick which results in excellent water permeability but defects
  during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of carbon
  nanotubes show promise \cite{humplik_nanostructured_2011} but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Zeolites have sub-nm pores
  with a narrow pore size distribution and MD simulations exhibit complete
  rejection of solvated ions \cite{murad_molecular_1998}, however, experimental
  rejection was low and attributed to interstitial defects formed during membrane
  synthesis \cite{li_desalination_2004}. There is a need for a scalable
  nanostructured membrane.

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production \cite{feng_scalable_2014}. LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied
  experimentally
  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h} phase
  (Fig.\ref{fig:assembly}). The presence of ca. 10\% wt water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations \cite{zhou_supported_2005}.  One can
  envision tailoring the pore region for specific separations by changing the
  monomer chemistry \cite{resel_h2-phase_2000}.

  Research into LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that the same result could be obtained using a second technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  %BJC: TODO: Scalable vector graphic for monomer structure, make bigger. Convert .svg to .eps
  %           Thicken slabs
  %           Make pore region on slab (blue) bigger
  %           Bump up atom sizes in atomistic render
  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The liquid crystal monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnar mesophases (e).}~\label{fig:assembly}
  \end{figure}

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al.~varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  In a later
  study of rejection performance, it was shown that membranes formed by Na-GA3C11
  cannot perform separations of solutes less than 1.2 nm in diameter because the
  pores are too large \cite{zhou_supported_2005}.  We do not yet understand how
  to controllably reduce the effective pore size or how to tune the chemical
  environment in the nanopores for effective water desalination and small organic
  molecule separations. The only source of predictive modeling for LLC systems
  have been macroscopic models which likely do not adequately describe transport
  at these length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from X-ray diffraction (XRD) experiments
  and in matching ionic conductivity measurements \cite{feng_thin_2016}.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe solute transport and
  suggest governing mechanisms. We can observe how the choice of head group may
  influence pore size for size exclusion driven separations. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential which affects the degree to which the membrane can exclude
  charged species.

  % BJC: TODO: Make new rendering of wedges
  %            Change color scheme 
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{Our previous level of understanding of the pore structure (a) only allows us to speculate
	   about transport behavior in the pore. A detailed molecular model (b) will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured from atomistic simulations. The hydrophilic region is in red and the 
	   hydrophobic region is colored blue.}~\label{fig:detail}
  \end{figure}

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data. The two
  phases appear to share similar structural characteristics since the pore 
  spacings in each system are in close agreement \cite{feng_thin_2016,resel_structural_2000}.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically
  requires many atoms which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

  \item What constitutes a pore and how well-defined are the pore regions? \label{point:poredefinition}

>>>>

  Nanostructured porous membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control pore architecture at the molecular scale,
  thereby permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

  Current commercial RO and NF membranes suffer from limitations inherent to
  their fabrication. Although scalable, each of these two types of membranes has
  a degree of stochasticity that makes overcoming the permeability-selectivity
  tradeoff a challenge.  Namely, it is difficult to increase the permeability of
  a desired molecular or atomic species, while maintaining the same retention of
  an undesired species \cite{werber_materials_2016}.

@@

  Commercial RO membranes are typically dense thin-film composite (TFC)
  membranes with a porous support layer and an active layer made of a polymer
  matrix formed through interfacial polymerization \cite{jeong_interfacial_2007}.
  During interfacial polymerization, a step-growth polymerization reaction occurs
  at the interface of an aqueous monomer solution that has been soaked into the
  support layer, and an organic solution of a second monomer. The polymer matrix
  is dense and highly cross-linked which gives high rejections but low
  permeability.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically porous membranes made by a phase-inversion
  process.  The most widely used phase-inversion process is immersion
  precipitation, during which a polymer dissolved in a solvent is submerged in a
  non-solvent. A solid, porous polymer membrane is all that remains once all
  solvent has been removed by non-solvent exchange
  \cite{smolders_microstructures_1992}.  The resultant pores are polydisperse in
  size, which hurts membrane selectivity. A second technique used to create NF
  membranes is call track-etching in which a polymer film is bombarded with
  charged particles, then chemically etched to create pores
  \cite{apel_track_2001}. The pores are uniform, which benefits selectivity;
  however, the membranes have a low porosity and subsequently low permeability.

@@

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-generation nanoporous
  membranes with high selectivity permitted by a precisely controlled pore size
  and high permeability allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Development of nanostructured porous materials has been limited by the
  ability to synthesize and scale various fundamentally sound technologies.
  Graphene sheets are atomically thick which results in excellent water
  permeability but defects during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of
  carbon nanotubes show promise \cite{humplik_nanostructured_2011}, but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Zeolites have sub-nm
  pores with a narrow pore size distribution, and MD simulations of these
  materials show that they exhibit complete rejection of solvated ions
  \cite{murad_molecular_1998}- however, experimental rejection was low and
  attributed to interstitial defects formed during membrane synthesis
  \cite{li_desalination_2004}. Consequently, there is a need for a scalable
  nanostructured porous membrane.

  Polymeric membranes based on self-assembling lyotropic liquid crystals (LLCs)
  are suitable candidates for aqueous separation applications. LLCs are
  amphiphilic molecules that share the characteristic ability of nanostructured
  porous membrane materials to create highly ordered nanostructures with the
  added benefits of low cost and synthetic techniques feasible for large-scale
  production \cite{feng_scalable_2014}. LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally
  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
  The neat Na-GA3C11 monomer forms the thermotropic liquid crystalline (TLC)
  Col\textsubscript{h} phase (Fig.\ref{fig:assembly}). The presence of ca. 10
  wt\% added water results in formation of the lyotropic H\textsubscript{II}
  phase. In both cases, Na-GA3C11 monomers assemble into mesophases made of
  hexagonally packed, uniform size, cylinders with hydrophilic head groups
  oriented inward towards the cylinder center. This LLC assembly is then
  polymerized in situ to form a cross-linked polymer network to stabilize the
  structure. The hydrophilic region can act as a pore for aqueous separations
  \cite{zhou_supported_2005}.  One can envision tailoring the pore region for
  specific separations by changing the monomer chemistry
  \cite{resel_h2-phase_2000}.

  Research into polymerized LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that the same result could be obtained using a second technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  %BJC: TODO: Scalable vector graphic for monomer structure, make bigger. Convert .svg to .eps
  %           Thicken slabs
  %           Make pore region on slab (blue) bigger
  %           Bump up atom sizes in atomistic render
  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The LLC monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnar mesophases (e).}~\label{fig:assembly}
  \end{figure}

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications.  Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}.  In a later study of rejection performance, it
  was shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} cannot perform separations of solutes less than 1.2 nm in
  diameter because the pores are too large \cite{zhou_supported_2005}.  We do not
  yet understand how to controllably reduce the effective pore size or how to
  tune the chemical environment in the nanopores of this or related materials for
  effective water desalination and small organic molecule separations. The only
  source of predictive modeling for LLC systems have been macroscopic models that
  likely do not adequately describe transport at these length scales
  \cite{hatakeyama_water_2011}. It will be challenging to efficiently narrow down
  the large design space in a laboratory setting without a robust model.

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from small angle X-ray scattering (SAXS)
  and wide angle X-ray scattering (WAXS) experiments as well as in matching ionic
  conductivity measurements \cite{feng_thin_2016}.

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group may influence pore size for size exclusion driven separations. We
  can interchange counterions which may influence both the pore size and the
  strength of the Donnan potential which affects the degree to which the membrane
  can exclude charged species.

  % BJC: TODO: Make new rendering of wedges
  %            Change color scheme 
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{Our previous level of understanding of the pore structure (a) only allows us to speculate
	   about transport behavior in the pore. A detailed molecular model (b) will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 NA-GA3C11 monomers
           are pictured from atomistic simulations. The hydrophilic region is in red and the 
	   hydrophobic region is colored blue.}~\label{fig:detail}
  \end{figure}

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport in these ordered, nanoporous
  organic systems, we must first gain a thorough understanding of the nanoscopic
  structure of LLC polymer membranes. Our approach to constructing a general
  model will follow the development of a model of the assembly formed by
  Na-GA3C11 since it has sufficient experimental characterization. We have also
  narrowed our scope to the development of a model of the Col\textsubscript{h}
  phase membrane. Compared to the H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point, due to the absence of
  water, and has detailed experimental wide-angle X-ray scattering (WAXS)
  patterns useful for reconstructing structural data. The two phases appear to
  share similar structural characteristics since the pore spacings in each system
  are in close agreement \cite{feng_thin_2016,resel_structural_2000}.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC phase. Simulating a membrane atomistically
  requires many atoms which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

  \item What constitutes a pore, and how well-defined are the pore regions? \label{point:poredefinition}
